Self-test prime dimensional EPR pairs with constant alphabet

### 2019-03-14 Draft/draft.tex
Honghao: fix a typo
Honghao: find another problem of the self-test proof

\textbf{The ideal strategy for the weighted $\CHSH_X$ test}.
Alice and Bob share state $\ket{\psi} = \frac{1}{\sqrt{d-1}}\sum_{i=1}^{d-1} \ket{ii}$ and choose the special subspace to be 
$V = \spn(\ket{1} ,\ket{d-1})$.
Alice's projectors and observables for the weighted $\CHSH_X$ test are
\begin{align*}
	&A_0^\diamond = \ketbra{1}{1} + \ketbra{d-1}{d-1}, &&A_0^\perp = \1 - A_\ast^\diamond, \\
	&A_{m_r+2} = A_{m_r+2}^0-A_{m_r+2}^1 = \ketbra{1}{1} - \ketbra{d-1}{d-1}, &&A_{m_r+2}^\perp = A_\ast^\perp,\\
	&A_{m_r+3} = A_{m_r+3}^0-A_{m_r+3}^1 = \ketbra{1}{d-1} + \ketbra{d-1}{1} ,&&A_{m_r+3}^\perp = A_\ast^\perp.
\end{align*}
Recall that in the extended weighted $\CHSH_X$ test, what matters is the 
Bob's observables restricted to the subspace $V$, which are given by
\begin{align*}
	B_1|_V = \cos(\frac{\pi}{d})(\ketbra{1}{1} - \ketbra{d-1}{d-1}) - \sin(\frac{\pi}{d})(\ketbra{1}{d-1} + \ketbra{d-1}{1}),\\
	B_2|_V = \cos(\frac{\pi}{d})(\ketbra{1}{1} - \ketbra{d-1}{d-1}) +\sin(\frac{\pi}{d})(\ketbra{1}{d-1} + \ketbra{d-1}{1}).
\end{align*}
To make sure operator $B_2$ commutes with the operator $U = \Psi(x_3x_4) = \Psi(u)$ that we will construct later, we pick a 
special basis of $X|_V =B_1|_VB_2|_V$, which is
\begin{align}
	\ket{x_1} = \frac{-1}{\sqrt{2}}(\ket{1} + i\ket{d-1}), &&
	\ket{x_{d-1}} = \frac{-e^{i\pi/d}}{\sqrt{2}}(\ket{1} - i\ket{d-1}).
\end{align}
It can be checked that $X\ket{x_1} = \omega_d \ket{x_1}$ and $X\ket{x_{d-1}} = \omega_d^{d-1} \ket{x_{d-1}}$.
In this basis, $B_1|_V$ and $B_2|_V$ are expressed as 
\begin{align}
	&B_1|_V = \omega_d\ketbra{x_1}{x_{d-1}} + \omega_d^{-1} \ketbra{x_{d-1}}{x_1},\\
	&B_2|_V =\ketbra{x_1}{x_{d-1}}  + \ketbra{x_{d-1}}{x_1}.
\end{align}

@@

Suppose $A_{m_r+1} = A_\ast^\diamond- A_\ast^\perp = \Pi_{V_A} - \Pi_{V_A^\perp}$m, where $V_A$ is a
$2m$-dimensional vector space and $\Pi_{V_A}$ is the projector onto it. The reason why it has dimension $2m$
will be clear shortly. On Bob's side, we also have $B_{n_r+1} = B_\ast^\diamond- B_\ast^\perp = \Pi_{V_B} - \Pi_{V_B^\perp}$.
Recall that from the extended weighted CHSH test, we know
\begin{align}
	A_\ast^\diamond \ket{\psi} = B_\ast^\diamond \ket{\psi} = A_\ast^\diamond \x B_\ast^\diamond \ket{\psi},
\end{align}
which implies that with an appropriate change of basis, we can get $V_A = V_B$, so in the rest of the proof
we drop the subscript of $V$.
By \cref{lm:chsh_comp} and \cref{prop:2d-subspace}, we know $V$ consists of $\omega_d$-eigenvectors and $\omega_d^{-1}$ eigenvectors of 
$X_A$, so we can write 
\begin{align}
	\Pi_{V} = \Pi_{V_1} + \Pi_{V_{d-1}},
\end{align}
where $\Pi_{V_{1}}$ is the projector onto the $\omega_d$-eigenspace of $X_A$ and $\Pi_{V_{d-1}}$ is the projector 
onto the $\omega_d^{-1}$-eigenspace of $X_A$.
Suppose $\ket{x_{1}} \in V_{1}$, then $X_A \ket{x_{1}} = \omega_d \ket{x_{1}}$.
By \cref{eq:ux_relation} we can calculate that
\begin{align}
\label{eq:ladder}
 X_AU_A^\dagger \ket{x_{1}} = U_A^\dagger X_A^r \ket{x_{1}} = \omega_d^r U_A^\dagger \ket{x_{1}},
\end{align}
so $U_A^\dagger \ket{x_{1}}$ is an eigenvector of $X_A$ with eigenvalue $\omega_d^r$.
By induction, we know $X_A (U_A^\dagger)^i \ket{x_{1}} = \omega_d^{r^i} (U_A^\dagger)^i\ket{x_{1}}$. 
From the set $\{(U_A^\dagger)^i \Pi_{V_1} (U_A)^i \}_{i=0}^{d-2}$, we can identify $\Pi_{V_i}$ for $i = 1 \dots  d-1$
such that $\Pi_{V_i}$ is the projector onto the $\omega_d^i$-eigenspace of $X_A$,
and 
\begin{align}
 \cup_{i \in [d-1]+1} V_i \subset \supp(\rho_A).
\end{align}
Since unitary transformation does not change the rank of a matrix, we know $\rank(\Pi_{V_1}) = \rank(\Pi_{V_{i}}) =N$
for $ i =1 \dots d-1$.
We pick a basis for $V_1$ such that 
\begin{align}
	\Pi_{V_1} =  \sum_{j=1}^m \ketbra{x_{1,j}}{x_{1,j}},
\end{align}
then we can construct 
\begin{align}
 \Pi_{V_i} = \sum_{j=1}^m \ketbra{x_{i,j}}{x_{i,j}},
\end{align}
where $\ket{x_{i,j}} = (U\ct)^{k_i} \ket{x_{1,j}}$ for $r^{k_i} \equiv i \pmod{d}$ and $1 \leq j \leq m$.
By \cref{eq:ux_relation} we also know that $U_A \ket{x_{i,j}} = \ket{x_{i/2,j}}$ for $i = 1,2 \dots d-1$.
In order to apply \cref{lm:ux_independ}, we construct $m$ subspaces $\{W_j\}_{j=1}^m$ where 
\begin{align}
	W_j = \spn( \{ \ket{x_{i,j}} \}_{i=1}^{d-1} )
\end{align}
The subspace $W_j$ is orthogonal to $W_{j'}$ for $j \neq j'$, and
$U_A$ and $X_A$ satisfy the condition of \cref{lm:ux_independ} when their actions are 
restricted to each $W_j$.
Similar argument also applies to operator $X_B$ and $U_B$ on Bob's side.

@@

\begin{proof}
The first step is to find a sum-of-square decomposition of 
the following Bell expression
\begin{align}
	\bar{\I}_\alpha = 2\sqrt{\alpha^2+1} \1 - \I_\alpha
	= \frac{2}{\sin(\mu)} \1 - \frac{\cos(\mu)}{\sin(\mu)}(A_1B_1+A_1B_2) -  A_2B_1 + A_2B_2.
\end{align} 
With the following notation
\begin{align*}
	Z_A = A_1 &\quad X_A = A_2\\
	Z_B = \frac{B_1+B_2}{2\cos(\mu)} &\quad X_B = \frac{B_1-B_2}{2\sin(\mu)},
\end{align*}
the two SOS decompositions that we use are
\begin{align}
	\label{eq:sos1}&\bar{\I}_\alpha = \frac{\sin(\mu)\bar{\I}_\alpha^2 + 4\sin(\mu)\cos(\mu)^2(Z_AX_B+X_AZ_B)^2}{4},\\
	\label{eq:sos2}&\bar{\I}_\alpha = \frac{\cos^2(\mu)}{\sin(\mu)}(Z_A-Z_B)^2 + \sin(\mu) (X_A-X_B)^2.
\end{align}
The verification is omitted here.

>>>>

\textbf{The ideal strategy for the weighted $\CHSH_X$ test}.
Alice and Bob share state $\ket{\psi} = \frac{1}{\sqrt{d-1}}\sum_{i=1}^{d-1} \ket{ii}$ and choose the special subspace to be 
$V = \spn(\ket{1} ,\ket{d-1})$.
Alice's projectors and observables for the weighted $\CHSH_X$ test are
\begin{align*}
	&A_0^\diamond = \ketbra{1}{1} + \ketbra{d-1}{d-1}, &&A_0^\perp = \1 - A_\ast^\diamond, \\
	&A_{m_r+2} = A_{m_r+2}^0-A_{m_r+2}^1 = \ketbra{1}{1} - \ketbra{d-1}{d-1}, &&A_{m_r+2}^\perp = A_\ast^\perp,\\
	&A_{m_r+3} = A_{m_r+3}^0-A_{m_r+3}^1 = \ketbra{1}{d-1} + \ketbra{d-1}{1} ,&&A_{m_r+3}^\perp = A_\ast^\perp.
\end{align*}
Recall that in the extended weighted $\CHSH_X$ test, what matters is the 
Bob's observables restricted to the subspace $V$, which are given by
\begin{align*}
	B_1|_V = \cos(\frac{\pi}{d})(\ketbra{1}{1} - \ketbra{d-1}{d-1}) - \sin(\frac{\pi}{d})(\ketbra{1}{d-1} + \ketbra{d-1}{1}),\\
	B_2|_V = \cos(\frac{\pi}{d})(\ketbra{1}{1} - \ketbra{d-1}{d-1}) +\sin(\frac{\pi}{d})(\ketbra{1}{d-1} + \ketbra{d-1}{1}).
\end{align*}
To make sure operator $B_2$ commutes with the operator $U = \Psi(x_3x_4) = \Psi(u)$ that we will construct later, we pick a 
special basis of $X|_V =B_1|_VB_2|_V$, which is
\begin{align}
	\ket{x_1} = \frac{-1}{\sqrt{2}}(\ket{1} + i\ket{d-1}), &&
	\ket{x_{d-1}} = \frac{-e^{i\pi/d}}{\sqrt{2}}(\ket{1} - i\ket{d-1}).
\end{align}
It can be checked that $X\ket{x_1} = \omega_d \ket{x_1}$ and $X\ket{x_{d-1}} = \omega_d^{d-1} \ket{x_{d-1}}$.
In this basis, $B_1|_V$ and $B_2|_V$ are expressed as 
\begin{align}
	&B_1|_V = \omega_d\ketbra{x_1}{x_{d-1}} + \omega_d^{-1} \ketbra{x_{d-1}}{x_1},\\
	&B_2|_V =\ketbra{x_1}{x_{d-1}}  + \ketbra{x_{d-1}}{x_1}.
\end{align}
\hl{If we express $\ket{j}$'s as $\ket{x_j}$'s, then we can find that 
shared state is}\\
\hl{ $\sum_{j=1}^{d-1} \ket{j}\ket{j} = \sum_{j=0}^{(d-1)/2} e^{-ij\pi/d} (\ket{x_{j}}\ket{x_{d-j}} + \ket{x_{d-j}}\ket{x_{j}}$.}

@@

Suppose $A_{m_r+1} = A_\ast^\diamond- A_\ast^\perp = \Pi_{V_A} - \Pi_{V_A^\perp}$m, where $V_A$ is a
$2m$-dimensional vector space and $\Pi_{V_A}$ is the projector onto it. The reason why it has dimension $2m$
will be clear shortly. On Bob's side, we also have $B_{n_r+1} = B_\ast^\diamond- B_\ast^\perp = \Pi_{V_B} - \Pi_{V_B^\perp}$.
Recall that from the extended weighted CHSH test, we know
\begin{align}
	A_\ast^\diamond \ket{\psi} = B_\ast^\diamond \ket{\psi} = A_\ast^\diamond \x B_\ast^\diamond \ket{\psi},
\end{align}
which implies that with an appropriate change of basis, we can get $V_A = V_B$, so in the rest of the proof
we drop the subscript of $V$.
By \cref{lm:chsh_comp} and \cref{prop:2d-subspace}, we know $V$ consists of $\omega_d$-eigenvectors and $\omega_d^{-1}$ eigenvectors of 
$X_A$, so we can write 
\begin{align}
	\Pi_{V} = \Pi_{V_1} + \Pi_{V_{d-1}},
\end{align}
where $\Pi_{V_{1}}$ is the projector onto the $\omega_d$-eigenspace of $X_A$ and $\Pi_{V_{d-1}}$ is the projector 
onto the $\omega_d^{-1}$-eigenspace of $X_A$.
\hl{Here what we need is that $\Pi_{V_1,A} \ket{\psi} = \Pi_{V_{d-1},B} \ket{\psi} = \Pi_{V_1,A}\Pi_{V_{d-1},B}\ket{\psi}$
and $\Pi_{V_1,B} \ket{\psi} = \Pi_{V_{d-1},A} \ket{\psi} = \Pi_{V_1,B}\Pi_{V_{d-1},A}\ket{\psi}$.
This is not proved because $\Pi_{V_1}$ and $\Pi_{V_{d-1}}$ are not in the strategy.}
Suppose $\ket{x_{1}} \in V_{1}$, then $X_A \ket{x_{1}} = \omega_d \ket{x_{1}}$.
By \cref{eq:ux_relation} we can calculate that
\begin{align}
\label{eq:ladder}
 X_AU_A^\dagger \ket{x_{1}} = U_A^\dagger X_A^r \ket{x_{1}} = \omega_d^r U_A^\dagger \ket{x_{1}},
\end{align}
so $U_A^\dagger \ket{x_{1}}$ is an eigenvector of $X_A$ with eigenvalue $\omega_d^r$.
By induction, we know $X_A (U_A^\dagger)^i \ket{x_{1}} = \omega_d^{r^i} (U_A^\dagger)^i\ket{x_{1}}$. 
From the set $\{(U_A^\dagger)^i \Pi_{V_1} (U_A)^i \}_{i=0}^{d-2}$, we can identify $\Pi_{V_i}$ for $i = 1 \dots  d-1$
such that $\Pi_{V_i}$ is the projector onto the $\omega_d^i$-eigenspace of $X_A$,
\hl{Actually, we are not sure about the containment between $\supp((U_A\ct)^{\log_r(d-1)} \Pi_{V_1} U_A^{\log_r(d-1)})$ and 
$\supp(\Pi_{V_{d-1}})$, so we cannot say that $\Pi_{V_{d-1}} = (U_A\ct)^{\log_r(d-1)} \Pi_{V_1} U_A^{\log_r(d-1)}$.}
and 
\begin{align}
 \cup_{i \in [d-1]+1} V_i \subset \supp(\rho_A).
\end{align}
Since unitary transformation does not change the rank of a matrix, we know $\rank(\Pi_{V_1}) = \rank(\Pi_{V_{i}}) =N$
for $ i =1 \dots d-1$.
We pick a basis for $V_1$ such that 
\begin{align}
	\Pi_{V_1} =  \sum_{j=1}^m \ketbra{x_{1,j}}{x_{1,j}},
\end{align}
then we can construct 
\begin{align}
 \Pi_{V_i} = \sum_{j=1}^m \ketbra{x_{i,j}}{x_{i,j}},
\end{align}
where $\ket{x_{i,j}} = (U\ct)^{k_i} \ket{x_{1,j}}$ for $r^{k_i} \equiv i \pmod{d}$ and $1 \leq j \leq m$.
By \cref{eq:ux_relation} we also know that $U_A \ket{x_{i,j}} = \ket{x_{i/2,j}}$ for $i = 1,2 \dots d-1$.
In order to apply \cref{lm:ux_independ}, we construct $m$ subspaces $\{W_j\}_{j=1}^m$ where 
\begin{align}
	W_j = \spn( \{ \ket{x_{i,j}} \}_{i=1}^{d-1} )
\end{align}
The subspace $W_j$ is orthogonal to $W_{j'}$ for $j \neq j'$, and
$U_A$ and $X_A$ satisfy the condition of \cref{lm:ux_independ} when their actions are 
restricted to each $W_j$.
Similar argument also applies to operator $X_B$ and $U_B$ on Bob's side.

@@

\begin{proof}
The first step is to find a sum-of-square decomposition of 
the following Bell expression
\begin{align}
	\bar{\I}_\alpha = 2\sqrt{\alpha^2+1} \1 - \I_\alpha
	= \frac{2}{\sin(\mu)} \1 - \frac{\cos(\mu)}{\sin(\mu)}(A_1B_1+A_1B_2) -  A_2B_1 + A_2B_2.
\end{align} 
With the following notation
\begin{align*}
	&Z_A = A_1 && X_A = A_2\\
	&Z_B = \frac{B_1+B_2}{2\cos(\mu)} && X_B = \frac{B_1-B_2}{2\sin(\mu)},
\end{align*}
the two SOS decompositions that we use are
\begin{align}
	\label{eq:sos1}&\bar{\I}_\alpha = \frac{\sin(\mu)\bar{\I}_\alpha^2 + 4\sin(\mu)\cos(\mu)^2(Z_AX_B+X_AZ_B)^2}{4},\\
	\label{eq:sos2}&\bar{\I}_\alpha = \frac{\cos^2(\mu)}{\sin(\mu)}(Z_A-Z_B)^2 + \sin(\mu) (X_A-X_B)^2.
\end{align}
The verification is omitted here.
